---
title: "MLwiN - life satisfaction"
author: "Sunyeop Lee"
date: "3/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

```{r load packages, include=F}
library(R2MLwiN);options(MLwiN_path = "MLwiN v3.04/mlwin.exe")
library(lme4)
library(lmtest)
library(texreg)
library(grid)
library(gridExtra)
library(tidyverse)
library(car)
library(ggrepel)
library(mice)

```
 
```{r call relevant R scripts, include=F}
source("data preparation.R")
```
# scatter plots with correlation
```{r}
# correlation: gdp vs. life satisfaction
{
grob1 <- grobTree(textGrob(paste("r = ", round(cor(mywvs6_country$gdp,
                                                   mywvs6_country$life_satisfaction), 4)),
                                 x = 0.6, y = 0.17, hjust = 0,
                                 gp = gpar(col = "black", fontsize = 9, fontface = "bold")))
grob2 <- grobTree(textGrob(paste("(p = ",round(cor.test(mywvs6_country$gdp,
                                                        mywvs6_country$life_satisfaction)$p.value,4),
                                 ")"), x = 0.6, y = 0.1, hjust = 0,
                           gp = gpar(col = "black", fontsize = 9, fontface = "bold")))
corrplot_gdp_ls <- ggplot(mywvs6_country, aes(x=gdp*10000, y=life_satisfaction)) + 
  geom_point() + 
  ggtitle("Life satisfaction by GDP") + 
  geom_smooth(method=lm, se=FALSE, color='black') + 
  scale_x_continuous(name = "GDP per capita, PPP", breaks = c(0,25000,50000,75000,100000)) +
  scale_y_continuous(name = "Life satisfaction") +
  annotation_custom(grob1) + 
  annotation_custom(grob2) + 
  theme(plot.title = element_text(hjust = 0.5), panel.background = element_blank(),
        axis.line = element_line(color="black"), 
        axis.line.x = element_line(color="black"))
corrplot_gdp_ls
}
# correlation: lngdp vs. life satisfaction
{
grob1 <- grobTree(textGrob(paste("r = ", round(cor(mywvs6_country$lngdp,
                                                   mywvs6_country$life_satisfaction), 4)),
                                 x = 0.6, y = 0.17, hjust = 0,
                                 gp = gpar(col = "black", fontsize = 9, fontface = "bold")))
grob2 <- grobTree(textGrob(paste("(p = ",round(cor.test(mywvs6_country$lngdp,
                                                        mywvs6_country$life_satisfaction)$p.value,4),
                                 ")"), x = 0.6, y = 0.1, hjust = 0,
                           gp = gpar(col = "black", fontsize = 9, fontface = "bold")))
corrplot_lngdp_ls <- ggplot(mywvs6_country, aes(x=gdp*10000, y=life_satisfaction)) + 
  geom_point() + 
  ggtitle("Life satisfaction by lnGDP") + 
  geom_smooth(method=lm, se=FALSE, color='black') + 
  scale_x_continuous(name = "GDP per capita, PPP", trans = "log", breaks = c(2500, 5000, 10000, 20000, 40000, 80000)) +
  scale_y_continuous(name = "Life satisfaction") +
  annotation_custom(grob1) + 
  annotation_custom(grob2) +
  theme(plot.title = element_text(hjust = 0.5), panel.background = element_blank(),
        axis.line = element_line(color="black"), 
        axis.line.x = element_line(color="black"))
corrplot_lngdp_ls
}
# correlation: gni vs. life satisfaction
{
grob1 <- grobTree(textGrob(paste("r = ", round(cor(mywvs6_country$gni,
                                                   mywvs6_country$life_satisfaction), 4)),
                                 x = 0.6, y = 0.17, hjust = 0,
                                 gp = gpar(col = "black", fontsize = 9, fontface = "bold")))
grob2 <- grobTree(textGrob(paste("(p = ",round(cor.test(mywvs6_country$gni,
                                                        mywvs6_country$life_satisfaction)$p.value,4),
                                 ")"), x = 0.6, y = 0.1, hjust = 0,
                           gp = gpar(col = "black", fontsize = 9, fontface = "bold")))
corrplot_gni_ls <- ggplot(mywvs6_country, aes(x=gni*10000, y=life_satisfaction)) + 
  geom_point() + 
  ggtitle("Life satisfaction by GNI") + 
  geom_smooth(method=lm, se=FALSE, color='black') + 
  scale_x_continuous(name = "GNI per capita, PPP", breaks = c(0,25000,50000,75000,100000)) +
  scale_y_continuous(name = "Life satisfaction") +
  annotation_custom(grob1) + 
  annotation_custom(grob2) +
  theme(plot.title = element_text(hjust = 0.5), panel.background = element_blank(),
        axis.line = element_line(color="black"), 
        axis.line.x = element_line(color="black"))
corrplot_gni_ls
}
# correlation: gini vs. life satisfaction
{
grob1 <- grobTree(textGrob(paste("r = ", round(cor(mywvs6_country$gini,
                                                   mywvs6_country$life_satisfaction), 4)),
                                 x = 0.6, y = 0.15, hjust = 0,
                                 gp = gpar(col = "black", fontsize = 9, fontface = "bold")))
grob2 <- grobTree(textGrob(paste("(p = ",round(cor.test(mywvs6_country$gini,
                                                        mywvs6_country$life_satisfaction)$p.value,4),
                                 ")"), x = 0.6, y = 0.08, hjust = 0,
                           gp = gpar(col = "black", fontsize = 9, fontface = "bold")))
corrplot_gini_ls <- ggplot(mywvs6_country, aes(x=gini, y=life_satisfaction)) + 
  geom_point() + 
  ggtitle("Life satisfaction by Gini") + 
  geom_smooth(method=lm, se=FALSE, color='black') + 
  scale_x_continuous(name = "Gini index", breaks = c(30,40,50,60)) +
  scale_y_continuous(name = "Life satisfaction") +
  annotation_custom(grob1) + 
  annotation_custom(grob2) +
  theme(plot.title = element_text(hjust = 0.5), panel.background = element_blank(),
        axis.line = element_line(color="black"), 
        axis.line.x = element_line(color="black"))
corrplot_gini_ls
}

grid.arrange(corrplot_gdp_ls,corrplot_gini_ls, nrow=1)

```


# unadjusted random intercept model - R2MLwiN - life satisfaction
```{r null random intercept model}
# unadjusted random intercept model
sample <- mywvs6
options <- list(debugmode = FALSE, Meth = 0, resi.store = TRUE,
                residoptions='standardised',
                weighting=list(weightvar=c(NA,"weight"), standardised=T))
model <-  life_satisfaction ~ 1 + (1 | country) + (1| id)

nullRI_ls <- runMLwiN(Formula = model, data = sample, estoption = options) # complete case analysis
summary(nullRI_ls)

# mi_nullRI_ls_gini <- lapply(mi_complete_final, function(x){ # analysis accounting for missing data
#   runMLwiN(Formula = model, data = x, estoption = options)
# })
# 
# combined_fit <- pool(mi_nullRI_ls_gini)
# summary(combined_fit)

# exploration and diagnostics
{
# # variance partitioning coefficient (i.e., intraclass correlation coefficient)
# vpc <- nullRI_ls["RP"][["RP2_var_Intercept"]]*100 / 
#           (nullRI_ls["RP"][["RP1_var_Intercept"]] + nullRI_ls["RP"][["RP2_var_Intercept"]])
# vpc

# # caterpillar plot for level 2 residuals
# residuals <- nullRI_ls@residual$lev_2_resi_est_Intercept
# residualsCI <- 1.96 * sqrt(nullRI_ls@residual$lev_2_resi_var_Intercept)
# residualsRank <- rank(residuals)
# rankno <- order(residualsRank)
# caterpillar(y = residuals[rankno], x = 1:60, qtlow = (residuals - residualsCI)[rankno],
#             qtup = (residuals + residualsCI)[rankno], xlab = 'Rank', ylab = 'Intercept')


# # normality diagnostic for level 2
# residuals <- nullRI_ls@residual$lev_2_resi_est_Intercept
# qqnorm(residuals, ylab="Standardized Residuals", xlab="Normal Scores", 
#          main="Life satisfaction level 2")
# qqline(residuals)
# 
# # normality diagnostic for level 1
# residuals <- nullRI_ls@residual$lev_1_resi_est_Intercept
# qqnorm(residuals, ylab="Standardized Residuals", xlab="Normal Scores", 
#          main="Life satisfaction level 1")
# qqline(residuals)
}
  
```

# Adjusted random intercept model - R2MLwiN - life satisfaction
```{r adjusted random intercept model}
# adjusted random intercept model
sample <- mywvs6
options <- list(debugmode = F, Meth = 0, resi.store = T, residoptions='standardised',
                weighting=list(weightvar=c(NA,"weight"), standardised=T))
model <-
  life_satisfaction ~ 1 + sex + age_c + age_c_2 + income + education + socialclass + maritalstatus + gdp_c + gini_c +
  (1 | country) + (1 | id)
fullRI_ls_gini <- runMLwiN(Formula = model, data = sample, estoption = options) # complete case analysis
summary(fullRI_ls_gini)

# mi_fullRI_ls_gini <- lapply(mi_complete_final, function(x){ # analysis accounting for missing data
#   runMLwiN(Formula = model, data = x, estoption = options)
# })
# combined_fit <- pool(mi_fullRI_ls_gini)
# summary(combined_fit)

# exploration and diagnostics
{
# # variance partitioning coefficient (i.e., intraclass correlation coefficient)
# vpc <- fullRI_ls["RP"][["RP2_var_Intercept"]]*100 / 
#           (fullRI_ls["RP"][["RP1_var_Intercept"]] + fullRI_ls["RP"][["RP2_var_Intercept"]])
# vpc

# # caterpillar plot for level 2 residuals
# residuals <- fullRI_ls@residual$lev_2_resi_est_Intercept
# residualsCI <- 1.96 * sqrt(fullRI_ls@residual$lev_2_resi_var_Intercept)
# residualsRank <- rank(residuals)
# rankno <- order(residualsRank)
# caterpillar(y = residuals[rankno], x = 1:60, qtlow = (residuals - residualsCI)[rankno],
#             qtup = (residuals + residualsCI)[rankno], xlab = 'Rank', ylab = 'Intercept')
# 
# # plot level 2 residual by GDP
# plot(x=(unique(mywvs6$gdp_c)+mean(mywvs6$gdp, na.rm=T))*10000,
#      y=fullRI_ls@residual$lev_2_resi_est_Intercept,
#      xlab="GDP ($)",
#      ylab="Level 2 residuals",
#      main="Life Satisfaction (level 2 residual) vs. GDP")



# # normality diagnostic for level 2
# residuals <- fullRI_ls@residual$lev_2_resi_est_Intercept
# qqnorm(residuals, ylab="Standardized Residuals", xlab="Normal Scores", 
#          main="Life satisfaction level 2")
# qqline(residuals)
# 
# hist(residuals) # histogram
# 
# # normality diagnostic for level 1
# residuals <- fullRI_ls@residual$lev_1_resi_est_Intercept
# qqnorm(residuals, ylab="Standardized Residuals", xlab="Normal Scores", 
#          main="Life satisfaction level 1")
# qqline(residuals)
# 
# hist(residuals)
# ggplot(mywvs6, aes(x=residuals)) + geom_histogram() # historgram
}
  
# # adjusted random intercept model - gini as tertiles
# sample <- mywvs6
# options <- list(debugmode = F, Meth = 0, resi.store = T, residoptions='standardised',
#                 weighting=list(weightvar=c(NA,"weight"), standardised=T))
# model <-
#   life_satisfaction ~ 1 + sex + age_c + age_c_2 + income + education + socialclass +
#   maritalstatus + lngdp + gini_2nd + gini_3rd +
#   (1 | country) + (1 | id)
# fullRI_ls_gini_tertile <- runMLwiN(Formula = model, data = sample, estoption = options) # complete case analysis
# summary(fullRI_ls_gini_tertile)

```

# unadjusted complex level 2 model - R2MLwiN - life satisfaction
```{r}
# unadjusted complex level 2 model
sample <- mywvs6
options <- list(debugmode = F, Meth = 0, resi.store = T, residoptions='standardised',
                weighting=list(weightvar=c(NA,"weight"), standardised=T), reset=c(2,2))
model <-  life_satisfaction ~ 1 + gini_c +
  (1 + gini_c| country) + (1| id)
nullRS_ls_gini <- runMLwiN(Formula = model, data = sample, estoption = options) # complete case analysis
summary(nullRS_ls_gini)

# mi_nullRS_ls_gini <- lapply(mi_complete_final, function(x){ # analysis accounting for missing data
#   runMLwiN(Formula = model, data = x, estoption = options)
# })
# combined_fit <- pool(mi_nullRS_ls_gini)
# summary(combined_fit)

# plot variance as a quadratic function of gini
{
sample$lev2variance <- nullRS_ls_gini["RP"][["RP2_var_Intercept"]] +
                    2 * nullRS_ls_gini["RP"][["RP2_cov_Intercept_gini_c"]] *
                    sample[["gini_c"]] + nullRS_ls_gini["RP"][["RP2_var_gini_c"]] * sample[["gini_c"]]^2

variance_only_null <- ggplot(data = sample, aes(gini, lev2variance)) +
                    geom_line(linetype = "solid", size = 1) +
                    labs(title="Variance by Gini",
                         x="Gini",
                         y = "Variance") +
                    theme(plot.title = element_text(size = 10, hjust = 0.5,
                                                    face = "bold"),
                          axis.text=element_text(size=10, face="bold"),
                          axis.title=element_text(size=10,face="bold"))
variance_only_null
}
# plot the predicted value and its 95% coverage bounds
{
u0 <- nullRS_ls_gini@residual[["lev_2_resi_est_Intercept"]]
u1 <- nullRS_ls_gini@residual[["lev_2_resi_est_gini_c"]]
b0 <- nullRS_ls_gini@FP[["FP_Intercept"]]
b1 <- nullRS_ls_gini@FP[["FP_gini_c"]]
mywvs6_lv2 <- data.frame(u0,u1,b0,b1)
mywvs6_lv2 <- cbind(country_gini, mywvs6_lv2)
mywvs6_lv2$country_ls <- b0 + u0 + (b1 + u1) * mywvs6_lv2$gini_c
mywvs6_lv2$lev2variance <- nullRS_ls_gini["RP"][["RP2_var_Intercept"]] +
                    2 * nullRS_ls_gini["RP"][["RP2_cov_Intercept_gini_c"]] *
                    mywvs6_lv2[["gini_c"]] + nullRS_ls_gini["RP"][["RP2_var_gini_c"]] *
  mywvs6_lv2[["gini_c"]]^2
mywvs6_lv2$fit <- b0 + b1 * mywvs6_lv2$gini_c
mywvs6_lv2$upperCI <- mywvs6_lv2$fit + 1.96*sqrt(mywvs6_lv2$lev2variance)
mywvs6_lv2$lowerCI <- mywvs6_lv2$fit - 1.96*sqrt(mywvs6_lv2$lev2variance)
## plot
plot(x=mywvs6_lv2$gini, y=mywvs6_lv2$country_ls)
mean_variance_null <- ggplot(data = mywvs6_lv2, aes(x=gini, y=fit)) +
  geom_point(aes(x=gini, y=country_ls)) +
  geom_line(size=1) +
  geom_ribbon(aes(ymin=lowerCI, ymax=upperCI),
              fill="blue", alpha=0.15) +
  #geom_text_repel(aes(x=gini, y=country_ls, label=country), color="black", size=4) +
  labs(title="Mean and coverage bounds", x="Gini index", y = "Life satisfaction") +
  theme(plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
        axis.text=element_text(size=10, face="bold"),
        axis.title=element_text(size=10,face="bold"))
mean_variance_null
}

# unadjusted complex level 2 model - gini as tertiles
sample <- mywvs6
diag <-  matrix(, nrow = 2, ncol = 1)
diag[1, 1] <-  2
diag[2, 1] <-  1
options <- list(debugmode = F, Meth = 0, resi.store = T, residoptions='standardised',
                weighting=list(weightvar=c(NA,"weight"), standardised=T), reset=c(2,2),
                smat=diag)
model <-  life_satisfaction ~ 1 + gini_2nd + gini_3rd +
  (gini_1st + gini_2nd + gini_3rd| country) + (1| id)
nullRS_ls_gini_tertile <- runMLwiN(Formula = model, data = sample, estoption = options) # complete case analysis
summary(nullRS_ls_gini_tertile)

# mi_nullRS_ls_gini_tertile <- lapply(mi_complete_final, function(x){ # analysis accounting for missing data
#   runMLwiN(Formula = model, data = x, estoption = options)
# })
# combined_fit <- pool(mi_nullRS_ls_gini_tertile)
# summary(combined_fit)

# bar plot
{
lv2variance <- c(nullRS_ls_gini_tertile@RP[["RP2_var_gini_1st1"]],
                 nullRS_ls_gini_tertile@RP[["RP2_var_gini_2nd1"]],
                 nullRS_ls_gini_tertile@RP[["RP2_var_gini_3rd1"]])
gini_tertile <- c('gini_1st', 'gini_2nd', 'gini_3rd')
mywvs6_lv2 <- tibble(gini_tertile, lv2variance)

barplot1 <- ggplot(mywvs6_lv2) +
  geom_bar(stat = 'identity',
           aes(x = gini_tertile, y = lv2variance),
           fill = 'orange') +
  scale_y_continuous("Variance", limits=c(0,1)) +
  labs(title = "Variance of country-level life satisfaction", x = "Gini tertile")
barplot1
}

# mean error plots 
{
u0 <- nullRS_ls_gini_tertile@residual[["lev_2_resi_est_gini_1st1"]]
u1 <- nullRS_ls_gini_tertile@residual[["lev_2_resi_est_gini_2nd1"]]
u2 <- nullRS_ls_gini_tertile@residual[["lev_2_resi_est_gini_3rd1"]]
b0 <- nullRS_ls_gini_tertile@FP[["FP_Intercept"]]
b1 <- nullRS_ls_gini_tertile@FP[["FP_gini_2nd1"]]
b2 <- nullRS_ls_gini_tertile@FP[["FP_gini_3rd1"]]
mywvs6_lv2 <- data.frame(u0,u1,u2,b0,b1,b2)
mywvs6_lv2 <- cbind(country_gini, mywvs6_lv2)

mywvs6_lv2$country_ls[mywvs6_lv2$gini_tertile=="[23.9,33.5]"] <-
  mywvs6_lv2$b0[mywvs6_lv2$gini_tertile=="[23.9,33.5]"] + 
  mywvs6_lv2$u0[mywvs6_lv2$gini_tertile=="[23.9,33.5]"]
mywvs6_lv2$country_ls[mywvs6_lv2$gini_tertile=="(33.5,40.2]"] <- 
  mywvs6_lv2$b0[mywvs6_lv2$gini_tertile=="(33.5,40.2]"] + 
  mywvs6_lv2$b1[mywvs6_lv2$gini_tertile=="(33.5,40.2]"] + 
  mywvs6_lv2$u1[mywvs6_lv2$gini_tertile=="(33.5,40.2]"]
mywvs6_lv2$country_ls[mywvs6_lv2$gini_tertile=="(40.2,59.4]"] <- 
  mywvs6_lv2$b0[mywvs6_lv2$gini_tertile=="(40.2,59.4]"] + 
  mywvs6_lv2$b2[mywvs6_lv2$gini_tertile=="(40.2,59.4]"] + 
  mywvs6_lv2$u2[mywvs6_lv2$gini_tertile=="(40.2,59.4]"]
mywvs6_lv2$country_ls_sd[mywvs6_lv2$gini_tertile=="[23.9,33.5]"] <-
  sqrt(nullRS_ls_gini_tertile@RP[["RP2_var_gini_1st1"]])
mywvs6_lv2$country_ls_sd[mywvs6_lv2$gini_tertile=="(33.5,40.2]"] <- 
  sqrt(nullRS_ls_gini_tertile@RP[["RP2_var_gini_2nd1"]])
mywvs6_lv2$country_ls_sd[mywvs6_lv2$gini_tertile=="(40.2,59.4]"] <- 
  sqrt(nullRS_ls_gini_tertile@RP[["RP2_var_gini_3rd1"]])

lv2variance <- c(nullRS_ls_gini_tertile@RP[["RP2_var_gini_1st1"]],
                 nullRS_ls_gini_tertile@RP[["RP2_var_gini_2nd1"]],
                 nullRS_ls_gini_tertile@RP[["RP2_var_gini_3rd1"]])
sd<- sqrt(lv2variance)

statsummary <- mywvs6_lv2 %>% 
  group_by(gini_tertile) %>% 
  summarize(mean=mean(country_ls, na.rm=T))
statsummary <- cbind(statsummary, sd)

crossbarplot1 <- ggplot(data=mywvs6_lv2, aes(x=gini_tertile, y=country_ls)) + 
  geom_jitter(color="mediumpurple", alpha=0.5, width=0.2) +
  geom_point(data=statsummary,aes(x=gini_tertile, y=mean), size=rel(2)) +
  geom_crossbar(data=statsummary,aes(x=gini_tertile, y=mean ,ymin=mean-1.96*sd,
                                     ymax=mean+1.96*sd), width=0.25, size=0.005) +
  #geom_text_repel(aes(label=country), color="black", size=4) +
  labs(x="Gini tertiles", y="Life satisfaction") +
  ggtitle("Life satisfaction by Gini tertiles") +
  theme_bw()
crossbarplot1
}

```

# adjusted complex level 2 model - R2MLwiN - life satisfaction
```{r}
# adjusted complex level 2 model
sample <- mywvs6
options <- list(debugmode = F, Meth = 0, resi.store = T, residoptions='standardised',
                weighting=list(weightvar=c(NA,"weight"), standardised=T), reset=c(2,2),
                maxiter=50)
model <-  life_satisfaction ~ 1 + sex + age_c + age_c_2 + income + education + socialclass + maritalstatus + gdp_c + gini_c + 
  (1 + gini_c| country) + (1| id)
fullRS_ls_gini <- runMLwiN(Formula = model, data = sample, estoption = options) # complete case analysis
summary(fullRS_ls_gini)

# plot variance as a quadratic function of gini
{
sample$lev2variance <- fullRS_ls_gini["RP"][["RP2_var_Intercept"]] +
                    2 * fullRS_ls_gini["RP"][["RP2_cov_Intercept_gini_c"]] *
                    sample[["gini_c"]] + fullRS_ls_gini["RP"][["RP2_var_gini_c"]] * sample[["gini_c"]]^2

variance_only_full <- ggplot(data = sample, aes(gini, lev2variance)) +
                    geom_line(linetype = "solid", size = 1) +
                    labs(title="Variance by Gini",
                         x="Gini",
                         y = "Variance") +
                    theme(plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
                          axis.text=element_text(size=10, face="bold"),
                          axis.title=element_text(size=10,face="bold"))
variance_only_full
}
# plot the predicted value and its 95% coverage bounds
{
## create a dataset at level 2 (level 2 mean, variance)
# level 2 mean and variance
u0 <- fullRS_ls_gini@residual[["lev_2_resi_est_Intercept"]]
u1 <- fullRS_ls_gini@residual[["lev_2_resi_est_gini_c"]]
b0 <- fullRS_ls_gini@FP[["FP_Intercept"]]
b1 <- fullRS_ls_gini@FP[["FP_gini_c"]]
mywvs6_lv2 <- data.frame(u0,u1,b0,b1)
mywvs6_lv2 <- cbind(country_gini, mywvs6_lv2)
mywvs6_lv2$country_ls <- b0 + u0 + (b1 + u1) * mywvs6_lv2$gini_c
mywvs6_lv2$lev2variance <- fullRS_ls_gini["RP"][["RP2_var_Intercept"]] +
                    2 * fullRS_ls_gini["RP"][["RP2_cov_Intercept_gini_c"]] *
                    mywvs6_lv2[["gini_c"]] + fullRS_ls_gini["RP"][["RP2_var_gini_c"]] *
  mywvs6_lv2[["gini_c"]]^2
mywvs6_lv2$fit <- b0 + b1 * mywvs6_lv2$gini_c
mywvs6_lv2$upperCI <- mywvs6_lv2$fit + 1.96*sqrt(mywvs6_lv2$lev2variance)
mywvs6_lv2$lowerCI <- mywvs6_lv2$fit - 1.96*sqrt(mywvs6_lv2$lev2variance)
# level 2 constant variance from random intercept model
mywvs6_lv2$lev2variance_constant <- fullRI_ls_gini["RP"][["RP2_var_Intercept"]]
mywvs6_lv2$upperCI_constant <- mywvs6_lv2$fit + 1.96*sqrt(mywvs6_lv2$lev2variance_constant)
mywvs6_lv2$lowerCI_constant <- mywvs6_lv2$fit - 1.96*sqrt(mywvs6_lv2$lev2variance_constant)
## plot
mean_variance_full<- ggplot(data = mywvs6_lv2, aes(x=gini, y=fit)) +
  geom_point(aes(x=gini, y=country_ls)) +
  geom_line(size=1) + 
  geom_ribbon(aes(ymin=lowerCI, ymax=upperCI), fill="gray10", alpha=0.15) +
  geom_line(aes(y=upperCI_constant), linetype="dashed") +
  geom_line(aes(y=lowerCI_constant), linetype="dashed") +
#  geom_text_repel(aes(x=gini, y=country_ls, label=country), color="black", size=4) +
  labs(title="Life satisfaction by Gini", x="Gini index", y = "Life satisfaction") +
  theme(plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
        axis.text=element_text(size=10, face="bold"),
        axis.title=element_text(size=10,face="bold"),
        panel.background=element_blank(),
        axis.line=element_line(color="black"),
        panel.grid.major=element_blank())
mean_variance_full
}

# grid.arrange(variance_only_null, mean_variance_null, variance_only_full, mean_variance_full, nrow=2, ncol=2)

lrtest(fullRI_ls_gini, fullRS_ls_gini)
linearHypothesis(fullRS_ls_gini, c("RP2_cov_Intercept_gini_c", "RP2_var_gini_c"))
linearHypothesis(fullRS_ls_gini, c("RP2_var_Intercept"))
linearHypothesis(fullRS_ls_gini, c("RP2_cov_Intercept_gini_c"))
linearHypothesis(fullRS_ls_gini, c("RP2_var_gini_c"))


# adjusted complex level 2 model - gini as tertile
sample <- mywvs6
diag <-  matrix(, nrow = 2, ncol = 1)
diag[1, 1] <-  2
diag[2, 1] <-  1
options <- list(debugmode = F, Meth = 0, resi.store = T, residoptions='standardised',
                weighting=list(weightvar=c(NA,"weight"), standardised=T), reset=c(2,2),
                smat=diag, maxiter=50)
model <-  life_satisfaction ~ 1 + sex + age_c + age_c_2 + income + education + socialclass + maritalstatus + gdp_c + gini_2nd + gini_3rd +
  (gini_1st + gini_2nd + gini_3rd| country) + (1| id)
fullRS_ls_gini_tertile <- runMLwiN(Formula = model, data = sample, estoption = options) # complete case analysis
summary(fullRS_ls_gini_tertile)

linearHypothesis(fullRS_ls_gini_tertile, "RP2_var_gini_1st1=RP2_var_gini_2nd1")
linearHypothesis(fullRS_ls_gini_tertile, "RP2_var_gini_2nd1=RP2_var_gini_3rd1")
linearHypothesis(fullRS_ls_gini_tertile, "RP2_var_gini_1st1=RP2_var_gini_3rd1")





# bar plot
{
lv2variance <- c(fullRS_ls_gini_tertile@RP[["RP2_var_gini_1st1"]],
                 fullRS_ls_gini_tertile@RP[["RP2_var_gini_2nd1"]],
                 fullRS_ls_gini_tertile@RP[["RP2_var_gini_3rd1"]])
gini_tertile <- c('gini_1st', 'gini_2nd', 'gini_3rd')
mywvs6_lv2 <- tibble(gini_tertile, lv2variance)

barplot2 <- ggplot(mywvs6_lv2) +
  geom_bar(stat = 'identity',
           aes(x = gini_tertile, y = lv2variance),
           fill = 'orange') +
  scale_y_continuous("Variance", limits=c(0,1)) +
  labs(title = "Variance of country-level life satisfaction", x = "Gini tertile")
barplot2
}

grid.arrange(barplot1, barplot2, nrow=1)

# mean error plots 
{
u0 <- fullRS_ls_gini_tertile@residual[["lev_2_resi_est_gini_1st1"]]
u1 <- fullRS_ls_gini_tertile@residual[["lev_2_resi_est_gini_2nd1"]]
u2 <- fullRS_ls_gini_tertile@residual[["lev_2_resi_est_gini_3rd1"]]
b0 <- fullRS_ls_gini_tertile@FP[["FP_Intercept"]]
b1 <- fullRS_ls_gini_tertile@FP[["FP_gini_2nd1"]]
b2 <- fullRS_ls_gini_tertile@FP[["FP_gini_3rd1"]]
mywvs6_lv2 <- data.frame(u0,u1,u2,b0,b1,b2)
mywvs6_lv2 <- cbind(country_gini, mywvs6_lv2)

mywvs6_lv2$country_ls[mywvs6_lv2$gini_tertile=="[23.9,33.5]"] <-
  mywvs6_lv2$b0[mywvs6_lv2$gini_tertile=="[23.9,33.5]"] + 
  mywvs6_lv2$u0[mywvs6_lv2$gini_tertile=="[23.9,33.5]"]
mywvs6_lv2$country_ls[mywvs6_lv2$gini_tertile=="(33.5,40.2]"] <- 
  mywvs6_lv2$b0[mywvs6_lv2$gini_tertile=="(33.5,40.2]"] + 
  mywvs6_lv2$b1[mywvs6_lv2$gini_tertile=="(33.5,40.2]"] + 
  mywvs6_lv2$u1[mywvs6_lv2$gini_tertile=="(33.5,40.2]"]
mywvs6_lv2$country_ls[mywvs6_lv2$gini_tertile=="(40.2,59.4]"] <- 
  mywvs6_lv2$b0[mywvs6_lv2$gini_tertile=="(40.2,59.4]"] + 
  mywvs6_lv2$b2[mywvs6_lv2$gini_tertile=="(40.2,59.4]"] + 
  mywvs6_lv2$u2[mywvs6_lv2$gini_tertile=="(40.2,59.4]"]

lv2variance <- c(fullRS_ls_gini_tertile@RP[["RP2_var_gini_1st1"]],
                 fullRS_ls_gini_tertile@RP[["RP2_var_gini_2nd1"]],
                 fullRS_ls_gini_tertile@RP[["RP2_var_gini_3rd1"]])
sd<- sqrt(lv2variance)

statsummary <- mywvs6_lv2 %>% 
  group_by(gini_tertile) %>% 
  summarize(mean=mean(country_ls, na.rm=T))
statsummary <- cbind(statsummary, sd)

crossbarplot2 <- ggplot(data=mywvs6_lv2, aes(x=gini_tertile, y=country_ls)) + 
  geom_point(color="black", alpha=1, position=position_jitter(width=0.2, seed=1)) +
  # geom_point(data=statsummary,aes(x=gini_tertile, y=mean), size=rel(2)) +
  geom_crossbar(data=statsummary,aes(x=gini_tertile, y=mean ,ymin=mean-1.96*sd,
                                     ymax=mean+1.96*sd), width=0.25, size=0.005) +
  geom_text_repel(aes(label=country), color="black", size=3,
  position=position_jitter(width=0.2, seed=1)) +
  labs(x=NULL, y="Life satisfaction") +
  ggtitle("Life satisfaction by Gini tertiles") +
  theme(plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
        axis.text=element_text(size=10, face="bold"),
        axis.text.x=element_text(size=10, face="bold", color="black", angle=5),
        axis.title=element_text(size=10,face="bold"),
        axis.line=element_line(color="black"),
        panel.grid.major=element_blank(),
        panel.background=element_blank()) +
  scale_x_discrete(labels=c("lower Gini","middle Gini","higher Gini"))
crossbarplot2
}

grid.arrange(crossbarplot1, crossbarplot2, nrow=1)

grid.arrange(mean_variance_full, crossbarplot2, nrow=1)


```

# final analyses on multiply imputed datasets
```{r}
# adjusted complex level 2 model
sample <- mywvs6
options <- list(debugmode = T, Meth = 0, resi.store = T, residoptions='standardised',
                weighting=list(weightvar=c(NA,"weight"), standardised=T), reset=c(2,2),
                maxiter=50)
model <-  life_satisfaction ~ 1 + sex + age_c + age_c_2 + income + education + socialclass + maritalstatus + gdp_c + gini_c + 
  (1 + gini_c| country) + (1| id)
# mi_fullRS_ls_gini <- lapply(mi_complete_final, function(x){ # analysis accounting for missing data ## not working
#   runMLwiN(Formula = model, data = x, estoption = options)
# })
combined_fit <- pool(mi_fullRS_ls_gini)
summary(combined_fit)

# adjusted complex level 2 model - gini as tertile
sample <- mywvs6
diag <-  matrix(, nrow = 2, ncol = 1)
diag[1, 1] <-  2
diag[2, 1] <-  1
options <- list(debugmode = F, Meth = 0, resi.store = T, residoptions='standardised',
                weighting=list(weightvar=c(NA,"weight"), standardised=T), reset=c(2,2),
                smat=diag, maxiter=50)
model <-  life_satisfaction ~ 1 + sex + age_c + age_c_2 + income + education + socialclass + maritalstatus + gdp_c + gini_2nd + gini_3rd +
  (gini_1st + gini_2nd + gini_3rd| country) + (1| id)
# mi_fullRS_ls_gini_tertile <- lapply(mi_complete_final, function(x){ # analysis accounting for missing data
#   runMLwiN(Formula = model, data = x, estoption = options)
# })
combined_fit <- pool(mi_fullRS_ls_gini_tertile)
summary(combined_fit)





```



# all models with MCMC - R2MLwiN - life satisfaction
```{r eval=F, echo=F}
# null random intercept model
# sample <- mywvs6
# options <- list(debugmode = F, EstM = 1, resi.store = T, mcmcMeth = list(burnin=500, iterations=5000))
# model <-  life_satisfaction ~ 1 + (1 | country) + (1| id)
# nullRI_ls_MCMC <- runMLwiN(Formula = model, data = sample, estoption = options)
# 
# summary(nullRI_ls_MCMC)
# 
# trajectories(nullRI_ls_MCMC)
# sixway2(nullRI_ls_MCMC@chains[, "RP2_var_Intercept", drop = FALSE], "RP2_var_Intercept")
# 
# # variance partitioning coefficient
# vpc_mcmc <- nullRI_ls_MCMC["chains"][,"RP2_var_Intercept"] /
#   (nullRI_ls_MCMC["chains"][,"RP1_var_Intercept"] +
#      nullRI_ls_MCMC["chains"][,"RP2_var_Intercept"])
# sixway2(vpc_mcmc, name="VPC_MCMC")

# # adjusted random intercept model
# sample <- mywvs6
# options <- list(debugmode = F, EstM = 1, resi.store = T, mcmcMeth = list(burnin=500, iterations=5000))
# model <-  life_satisfaction ~ 1 + sex + age_c + age_c_2 + income + education + 
#   socialclass +  maritalstatus + lngdp + gini_c + 
#   (1 | country) + (1 | id)
# fullRI_ls_MCMC <- runMLwiN(Formula = model, data = sample, estoption = options)
# 
# summary(fullRI_ls_MCMC)
# trajectories(fullRI_ls_MCMC)
# 
# sixway2(fullRI_ls_MCMC@chains[, "FP_gdp_c", drop = FALSE], "FP_gdp_c")
# sixway2(fullRI_ls_MCMC@chains[, "RP2_var_Intercept", drop = FALSE], "RP2_var_Intercept")
# 
# # variance partitioning coefficient
# vpc_mcmc <- fullRI_ls_MCMC["chains"][,"RP2_var_Intercept"] /
#   (fullRI_ls_MCMC["chains"][,"RP1_var_Intercept"] +
#      fullRI_ls_MCMC["chains"][,"RP2_var_Intercept"])
# sixway2(vpc_mcmc, name="VPC_MCMC")

# adjusted random slope model
sample <- mywvs6
options <- list(debugmode = F, EstM = 1, resi.store = T, residoptions='standardised',
                mcmcMeth = list(burnin=500, iterations=10000))
model <-  life_satisfaction ~ 1 + sex + age_c + age_c_2 + income_m + education_m + socialclass + maritalstatus_m + lngdp + gini_2nd + gini_3rd +
                              (gini_1st + gini_2nd + gini_3rd | country) + (1| id)
fullRS_ls_MCMC <- runMLwiN(Formula = model, data = sample, estoption = options)

summary(fullRS_ls_MCMC)
trajectories(fullRS_ls_MCMC)

# compare DIC
ComparingModels <- rbind(fullRI_ls_MCMC["BDIC"], fullRS_ls_MCMC["BDIC"])
rownames(ComparingModels) <- c("Random intercept", "Random slope")
ComparingModels

# sixway diagnostics
sixway2(fullRS_ls_MCMC@chains[, "FP_gini_2nd1", drop = FALSE], "FP_gini_2nd1")
sixway2(fullRS_ls_MCMC@chains[, "RP2_var_gini_1st1", drop = FALSE], "RP2_var_gini_1st1")
sixway2(fullRS_ls_MCMC@chains[, "RP2_var_gini_2nd1", drop = FALSE], "RP2_var_gini_2nd1")
sixway2(fullRS_ls_MCMC@chains[, "RP2_var_gini_3rd1", drop = FALSE], "RP2_var_gini_3rd1")

# plot variance as a quadratic function of gdp
sample$lev2variance <- fullRS_ls_MCMC["RP"][["RP2_var_Intercept"]] +
                    2 * fullRS_ls_MCMC["RP"][["RP2_cov_Intercept_gdp_c"]] *
                    sample[["gdp_c"]] + fullRS_ls_MCMC["RP"][["RP2_var_gdp_c"]] *
  sample[["gdp_c"]]^2 

ggplot(data = sample, aes(gdp*10000, lev2variance)) +
                    geom_line(linetype = "solid", size = 1) +
                    labs(title="Variance by GDP",
                         x="GDP",
                         y = "Variance") +
                    theme(plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
                          axis.text=element_text(size=10, face="bold"),
                          axis.title=element_text(size=10,face="bold"))

```

